### 2023-01-17 backend/fourier_figure.tex
Initial commit

%from various answers on https://tex.stackexchange.com/questions/127375/replicate-the-fourier-transform-time-frequency-domains-correspondence-illustrati

\documentclass[tikz]{standalone}
\usepackage{pgfplots}
\pgfplotsset{compat=1.8}
\usepackage{xcolor}

\usetikzlibrary{3d}
\input{head}

\begin{document}

\definecolor{lightgray}{rgb}{0.8, 0.8, 0.8} %grid of coordinate system, axes
\definecolor{midgray}{rgb}{0.6, 0.6, 0.6} %layer of border
\definecolor{darkgray}{rgb}{0.4, 0.4, 0.4} %curves, fill under curve

%x=(-18:0.9), y=(90:0.9), z=(42:1.1) <-- full 3D view
    \begin{tikzpicture}[scale=2.4] %create tikz picture

        \begin{axis}[ %create 3d plot within tikz
         set layers=standard, %use predefined layers
         view={30}{35}, %perspective adjustment %25 30
        domain=0:10, %plot limit in trait space direction
        samples y=1, %samples for dual space direction
        unit vector ratio*=1 2 1, %rescale unit vectors
        hide axis, %do not plot axes
        xtick=\empty, ytick=\empty, ztick=\empty, %no ticks on coordinate axes
        clip=false %let me plot outside the coordinate system
    ]
        %limit variables
        \def\xmax{100} %limits for curves and layers
        \def\xmin{0}
        \def\ymax{55}
        \def\ymin{5}
        \def\zmax{25}
        \def\zmin{-5}
        \def\xlayer{110} %dual space layer
        \def\sumcurve{0} %sum curve of fourier modes
        \def\amp{0.3}

        %fourier mode curves
        \pgfplotsinvokeforeach{1,2,3,4,5}{ %for each fourier mode

            \xdef\amp{0.08*(6-#1)}        
            \draw [<->,on layer=background, lightgray] (axis cs:0,#1,0) -- (axis cs:9.5,#1,0); %axes
            \addplot3 [on layer=main, blue, smooth, samples=200,domain=-0.5:9.5] (x-1,#1,{\amp*cos(100*#1/2*(x) - 50*#1)}); %plot curves

            \xdef\sumcurve{\sumcurve + \amp*cos(100*#1/2*(x) - 50*#1)} %add current curve to sumcurve
        }

        %transparent layers
        \fill[white,opacity=0.7] (\xmin,0,\zmin) -- (\xmin,0,\zmax) -- (\xmax,0,\zmax) -- (\xmax,0,\zmin) -- cycle; %transparent layer in trait space
        \fill[white,opacity=0.7] (\xlayer,\ymin,\zmin) -- (\xlayer,\ymin,\zmax) -- (\xlayer,\ymax,\zmax) -- (\xlayer,\ymax,\zmin) -- cycle; % transparent layer for dual space

        %grid lines
        \pgfplotsinvokeforeach{\xmin,\xmin+5,...,\xmax}{ %create horizontal grid lines (trait space layer)
            \draw[lightgray,opacity=0.6] (#1,0,\zmin) -- (#1,0,\zmax);
        }
        \pgfplotsinvokeforeach{\ymin,\ymin+2.5,...,\ymax}{ %create horizontal grid lines (dual space layer)
            \draw[lightgray,opacity=0.6] (\xlayer,#1,\zmin) -- (\xlayer,#1,\zmax);
        }
        \pgfplotsinvokeforeach{\zmin,\zmin+5,...,\zmax}{ %create vertical grid lines (both layers)
            \draw[lightgray,opacity=0.6] (\xmin,0,#1) -- (\xmax,0,#1);
            \draw[lightgray,opacity=0.6] (\xlayer,\ymin,#1) -- (\xlayer,\ymax,#1);
        }

        %borders
        \draw[midgray] (\xmin,0,\zmin) -- (\xmin,0,\zmax) -- (\xmax,0,\zmax) -- (\xmax,0,\zmin) -- cycle; %trait space layer borders
        \draw[midgray] (\xlayer,\ymin,\zmin) -- (\xlayer,\ymin,\zmax) -- (\xlayer,\ymax,\zmax) -- (\xlayer,\ymax,\zmin) -- cycle; %dual space layer borders

        %dots on dual space pane
        % \pgfplotsinvokeforeach{1,2,3,4,5}{ %for each fourier mode
        %         \node[circle,scale=1,fill=blue,on layer=main] at (axis cs:9.5,#1,0.4*5-0.4*#1) {}; %add dots on dual space pane
        % }

        %trait space curve
        \addplot3 [samples=1000,red,thick] (x-1,0,{\sumcurve});

    % dual space curve (for some reason removing this screws up the diagram, so I've just made it transparent)
    \addplot3 [samples=200,domain=0:5,green,opacity=0] (11.5,x,{2*sin((x-0.7)*180)^200*e^(-x/2)-0.3});

    %trait space domain axes and labels
    \begin{scope}
    \draw[<->,thick] (\xmin-5,0,\zmin) -- (\xmax+5,0,\zmin);
    \draw[->,thick] (0,0,\zmin) -- (0,0,\zmax+3);
    \node[scale=37, canvas is xz plane at y=0] at (0.45*\xmax,0,\zmin-8) {Trait Space $\mathcal{T}$};
    \node[scale=27, canvas is xz plane at y=0, rotate=90] at (\xmin-5,0,0.45*\zmax) {Density $\phi(x)$};
    \end{scope}

    %dual domain axes and labels
    \begin{scope}
    \draw[<->,thick] (\xlayer,\ymin,\zmin) -- (\xlayer,\ymax+4,\zmin);
    \draw[->,thick] (\xlayer,\ymax,\zmin) -- (\xlayer,\ymax,\zmax+3);
    \node[scale=20,canvas is yz plane at x=11.5] at (axis cs:0,3,-1.75) {Fourier dual space $\Tilde{\mathcal{T}}$};
    \node[scale=20,canvas is yz plane at x=11.5, rotate=90] at (axis cs:0,5.8,0.6) {Magnitude of $\phi_k$};
    \pgfplotsinvokeforeach{1,2,3,4,5}{
        \node[scale=20,blue,canvas is yz plane at x=11.5] at (axis cs:0,#1,-1.2) {$\phi_{#1}$}; %add labels

        \node[circle,scale=20,fill=blue,canvas is yz plane at x=11.5] at (axis cs:0,#1,0.4*5-0.4*#1) {};
    }
    \end{scope}
    \end{axis}
    \end{tikzpicture}

\end{document}