### 2023-02-04 ind_files/06_stochastic_discussion.tex
Separate results from discussion

\epigraph{\justifying At any moment, there exists only a narrow layer between the trivial and the impossible where mathematical discoveries are made. Therefore, an applied problem is either solved trivially or not solved at all. It is an altogether different story if an
applied problem is found to fit (or made to fit!) a new formalism interesting for a mathematician.}{Andrei Kolmogorov}

One striking feature that repeatedly shows up in our derivations is that finite populations exhibit phenomena that are not visible in infinite population models. 
Most notably, if we observe the change in trait frequencies instead of the change in densities, we saw that finite populations are subject to an additional evolutionary force that vanishes in infinite population models.