### 2023-03-14 backend/contributions.tex
Add CRediT contributions page as per bio dept guidelines

%----------------------------------------------
% Acknowledgements Header
\begin{center}
{\huge \bfseries Contributions}\\[1.5cm]
\end{center}
%----------------------------------------------
\centering
\begin{table}[ht!]
%\caption{}
%\begin{center}
	\begin{tabularx}{\textwidth}{ll}%   
\hline
{\bf Contributor Name} & {\bf Contributor Role} \\%
\hline
Ananda Shikhara Bhat & Conceptualization \\% formulation or evolution of overarching research goals and aims
                     & Methodology \\% Development or design of methodology; creation of models
                     & Formal analysis \\% Application of statistical, mathematical, computational, or other formal techniques to analyze or synthesize study data.
                     & Investigation \\% Conducting a research and investigation process, specifically performing the experiments, or data/evidence collection.
                     & Writing - Original Draft \\% Creation and/or presentation of the reported work, specifically writing the initial draft (including substantive translation).
                     & Writing - Review and Editing \\% Preparation, creation and/or presentation of the reported work by those from the original research group, specifically critical review, commentary or revision - including pre- or post-acceptance stages.
                     & Visualization \\% Preparation, creation and/or presentation of the reported work, specifically visualization/data presentation.
Vishwesha Guttal     & Conceptualization \\% formulation or evolution of overarching research goals and aims  
                     & Resources \\% Provision of study materials, reagents, materials, patients, laboratory samples, animals, instrumentation, computing resources, or other analysis tools.
                     & Supervision \\%	Oversight and leadership responsibility for the research activity planning and execution, including mentorship external to the core team.
Rohini Balakrishnan  & Supervision \\%	Oversight and leadership responsibility for the research activity planning and execution, including mentorship external to the core team.
\hline
\end{tabularx}
%\end{center}
\label{contributions_table}
\end{table}%
This contributor syntax is based on the Journal of Cell Science CRediT Taxonomy\footnote{\href{https://journals.biologists.com/jcs/pages/author-contributions}{https://journals.biologists.com/jcs/pages/author-contributions}}