### 2023-03-17 MS_thesis.tex
Rephrase abstract

\documentclass[twoside,12pt]{iiser-thesis-modified} %Use the custom .cls file we have with two sided paper and 12pt font

%Bio dept guidelines say the font should be in Arial (ew).
%Uncommenting out the below lines will do it, but I intend to keep them commented unless explicitly told to change font to Arial. Change compiler to XeLaTeX or LuaLaTeX for this to work
%\usepackage{fontspec}
%\setmainfont{Arial}

%%%%%%%%%%%%%%%%%%%

%My own list of packages and commands, for convenience
\input{head}

%%%%%%%%%%%%%%%%%%%

\usepackage{fancyhdr} %For fancy page headers
\setlength{\headsep}{1.8in} %add spacing between header and body text

%%%%%%%%%%%%%%%%%%%

%this file contains the bibliography
\addbibresource{refs.bib}

%%%%%%%%%%%%%%%%%%%
% Packages/Macros %
%%%%%%%%%%%%%%%%%%%
\usepackage{fullpage}
\setlength{\parskip}{1em}

%1.5 line spacing (as requested by IISER)
\usepackage{setspace}
\onehalfspacing

%Decide upto what depth the table of contents should list: subsection, subsubsection, etc
\setcounter{tocdepth}{3}

\catcode`\@=11
\catcode`\@=12

%%%%%%%%%%%%%%%%%%%
%  Abstract and   %
%  other details  %
%%%%%%%%%%%%%%%%%%%

\title{Eco-evolutionary dynamics of finite populations from first principles}
\author{Ananda Shikhara Bhat}

%Details of supervisors and experts
\supervisor{Vishwesha Guttal}
\department{Centre for Ecological Sciences, Indian Institute of Science} %department of supervisor
\cosupervisor{Rohini Balakrishnan}
\reader{Sutirth Dey} %Expert

\dedication{This thesis is dedicated to ?}

%Details of graduation date
\graduationyear{2023}
\academicyear{2022-2023}
\graduationmonth{May}

%The abstract goes here. IISER guidelines say it should be <= 250 words.
\thesisabstract{
Population biology is built on a strong mathematical foundation developed during the Modern Synthesis. Historically, many of these formalisms have often worked with infinite populations, ignoring the effects of demographic stochasticity. Finite population models in population and quantitative genetics usually assume a fixed population size and are of limited applicability in the real world, where population sizes routinely fluctuate. In this thesis, I analytically describe evolution in finite, fluctuating populations from biological first principles. Starting from a density-dependent `birth-death process’ describing an arbitrary closed population of individuals with discrete traits, I derive stochastic differential equations (SDEs) for how trait frequencies change over time. These SDEs reveal a new directional evolutionary force, `noise-induced selection’, that is particular to finite, fluctuating populations and has been largely overlooked in standard formulations of evolution. Noise-induced selection can cause deviations from neutrality and can reverse the direction of evolution predicted by infinite-population frameworks. My SDEs also recover well-known results such as the replicator-mutator equation and the Price equation in the infinite population limit and can help explain and organize several previous studies under a single theoretical framework. Finally, I extend the birth-death formalism to one-dimensional quantitative traits through a `stochastic field theory’ that yields frameworks such as Kimura's continuum-of-alleles and Lande's gradient dynamics in the infinite population limit. My work thus generalizes the formal structures of population biology to finite, fluctuating populations and predicts a new evolutionary force unique to such populations. I also discuss the implications for evolutionary agent-based models and empirical studies.
}

%The acknowledgements go here. IISER guidelines say it should be <= 250 words.
\acknowledgments{
Not more than 250 words
}
\allowdisplaybreaks %allow align envts to go across pages

%%%%%%%%%%%%%%%%%%%
%  The Document   %
%%%%%%%%%%%%%%%%%%%

\begin{document}

\pagestyle{plain} %Don't make the initial frontmatter fancy

%Add the frontmatter
\thesisfront

%wrap the declaration of every new \part in the \clear and the group to ensure that the fresh page created by \part does not have a page number.
% This hack-ey workaround is from https://tex.stackexchange.com/a/191189 and is needed because \part defines \pagestyle WITHIN the definition itself and so you cannot use \thispagestyle to mess with it
\cleardoublepage
\begingroup
\makeatletter
\let\ps@plain\ps@empty
\part{Motivation \& Outline}
\endgroup

%Implement the fancy headers for everything from here on out
\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{\thepage}
\fancyhead[CE]{\rule[-4ex]{0pt}{4ex}\footnotesize\itshape Eco-evolutionary dynamics of finite populations from first principles}
\fancyhead[CO]{\rule[-4ex]{0pt}{4ex}\footnotesize\itshape\nouppercase{\leftmark}}
\setlength{\headheight}{40pt} % as requested by fancyhdf
\renewcommand{\headrulewidth}{0pt}

\renewcommand{\chaptermark}[1]{\markboth{#1}{}} %gets rid of chapter number from header
\renewcommand{\sectionmark}[1]{\markright{#1}} %gets rid of section number from header

%From here on, we just need to import TeX files from the ind_files folder to include the actual content of the thesis

%Intro
\chapter{Introduction}
\input{ind_files/01_Introduction}

%The 'meat': Methods, results, theory, etc.
\cleardoublepage
\begingroup
\makeatletter
\let\ps@plain\ps@empty
\part{Theory}\label{part_theory}
\endgroup

\chapter{Mathematical background and an expository example}\label{chap_math_background}
\input{ind_files/02_math_background}

\chapter{Population dynamics from stochastic first principles}\label{chap_BD}
\input{ind_files/03_BD_processes}

\chapter{Stochastic field equations for the evolution of quantitative traits}\label{chap_infD_processes}
\input{ind_files/04_field_eqns}

%Discussion and summary
\cleardoublepage
\begingroup
\makeatletter
\let\ps@plain\ps@empty
\part{Major Takeaways \& Discussion}\label{part_summary}
\endgroup

\chapter{A unified view of population dynamics}\label{chap_unification}
\input{ind_files/05_unification_discussion}

\chapter{Discussion and Outlook}\label{sec_disc}
\input{ind_files/06_nontechnical_discussion}

%Add a new page after all the chapters are done, just to separate the main text from the appendices. If adding a new page results in you being on an even page, add another new page, just to ensure that you always begin with text on an odd page (useful if you intend to print the thesis at some point).
\clearpage
\null\thispagestyle{plain}\newpage
\ifodd\value{page}\else\null\thispagestyle{plain}\newpage\fi

%The \phantomsection is needed for hyperref to jump to the correct location
\cleardoublepage
\phantomsection
\appendix
\addcontentsline{toc}{part}{Appendices}

%make sure the title of the appendices say 'Appendix' and not 'Chapter' even though we declare them at `chapter level' internally
\renewcommand{\chaptername}{Appendix}

%For the headers of appendices (on odd pages), preface the appendix title with 'Appendix X', where X is the Appendix index (A,B,C,...)
\fancyhead[CO]{  \rule[-4ex]{0pt}{4ex}\footnotesize\itshape{\chaptername~\thechapter: \nouppercase{\leftmark}}}

%Import the Appendices
\chapter{From It\^{o} to Fokker-Planck}\label{App_SDE_FPE}
\input{ind_files/99_Appendix_A}
\chapter{Deriving stochastic trait frequency dynamics using It\^{o}'s formula}\label{App_density_to_freq}
\input{ind_files/99_Appendix_B}
\chapter{A Price-like equation for the variance of a type-level quantity}\label{App_stoch_var_eqns}
\input{ind_files/99_Appendix_C}
\chapter{Some Examples}\label{App_examples}
\input{ind_files/99_Appendix_D_examples}ne

%Bibliography
%Again, add the phantomsection so that hyperref knows where to go
\cleardoublepage
\phantomsection
\Huge\printbibheading
%Putting \printbibheading up here makes sure the heading is printed before I change the font size below.
%Change the font for the bibliography
\renewcommand*{\bibfont}{\footnotesize}	
%Add the bibliography
\printbibliography[heading=none]
%Add an entry in the table of contents
\addcontentsline{toc}{part}{References}

%And we're done :)
\end{document}