### 2023-03-22 MS_thesis.tex
Add acknowledgements

\documentclass[twoside,12pt]{iiser-thesis-modified} %Use the custom .cls file we have with two sided paper and 12pt font

%Bio dept guidelines say the font should be in Arial (ew).
%Uncommenting out the below lines will do it, but I intend to keep them commented unless explicitly told to change font to Arial. Change compiler to XeLaTeX or LuaLaTeX for this to work
%\usepackage{fontspec}
%\setmainfont{Arial}

%%%%%%%%%%%%%%%%%%%

%My own list of packages and commands, for convenience
\input{head}

%%%%%%%%%%%%%%%%%%%

\usepackage[final]{microtype} % microtype makes `subliminal improvements' in typesetting.
%see http://www.khirevich.com/latex/microtype/

\usepackage{fancyhdr} %For fancy page headers
\setlength{\headsep}{1.8in} %add spacing between header and body text

%%%%%%%%%%%%%%%%%%%

%this file contains the bibliography
\addbibresource{refs.bib}

%%%%%%%%%%%%%%%%%%%
% Packages/Macros %
%%%%%%%%%%%%%%%%%%%
\usepackage{fullpage}
\setlength{\parskip}{1em}

%1.5 line spacing (as requested by IISER)
\usepackage{setspace}
\onehalfspacing

%Decide upto what depth the table of contents should list: subsection, subsubsection, etc
\setcounter{tocdepth}{3}

\catcode`\@=11
\catcode`\@=12

%%%%%%%%%%%%%%%%%%%
%  Abstract and   %
%  other details  %
%%%%%%%%%%%%%%%%%%%

\title{Eco-evolutionary dynamics of finite populations from first principles}
\author{Ananda Shikhara Bhat}

%Details of supervisors and experts
\supervisor{Vishwesha Guttal}
\department{Centre for Ecological Sciences, Indian Institute of Science} %department of supervisor
\cosupervisor{Rohini Balakrishnan}
\reader{Sutirth Dey} %Expert

%Dedication
\dedication{This thesis is dedicated to ?}

%Details of graduation date
\graduationyear{2023}
\academicyear{2022-2023}
\graduationmonth{May}

%The abstract goes here. IISER guidelines say it should be <= 250 words.
\thesisabstract{
Population biology is built on a strong mathematical foundation developed during the Modern Synthesis. Historically, much of this foundation has been built using models of infinite populations, ignoring the effects of demographic stochasticity. Finite population models in population and quantitative genetics usually assume a fixed population size and are of limited applicability in a world where population sizes  routinely fluctuate. In this thesis, I analytically describe evolution in finite, fluctuating populations from biological first principles. Starting from a density-dependent `birth-death process’ describing an arbitrary closed population of individuals with discrete traits, I derive stochastic differential equations (SDEs) for how trait frequencies change over time. These SDEs reveal a new directional evolutionary force, `noise-induced selection’, that is particular to finite, fluctuating populations and has been largely overlooked in standard formulations of evolution. Noise-induced selection can cause deviations from neutrality and can reverse the direction of evolution predicted by infinite-population frameworks. My SDEs also recover well-known results such as the replicator-mutator equation and the Price equation in the infinite population limit and can help explain and organize several previous studies under a single theoretical framework. Finally, I extend the birth-death formalism to one-dimensional quantitative traits through a `stochastic field theory’ that yields frameworks such as Kimura's continuum-of-alleles and Lande's gradient dynamics in the infinite population limit. My work thus generalizes the formal structures of population biology to finite, fluctuating populations and predicts a new evolutionary force unique to such populations. I also discuss the implications for evolutionary agent-based models and empirical studies.
}

%The acknowledgements go here.
\acknowledgments{
%\vspace{-0.5cm}
First and foremost, I am grateful to my advisors, Prof. Vishwesha Guttal and Prof. Rohini Balakrishnan, for their masterful mentoring. The enormous academic freedom they gave me to pursue essentially whichever ideas seemed promising at the time, coupled with their readiness to provide detailed guidance whenever I asked for it, made working with them an incredibly enriching experience. Their constant focus on the bigger picture was vital to helping me see the forest when I was stuck looking at the trees. Both Vishu and Rohini are brilliant scientists and fantastic mentors, and their mentorship and guidance have deeply impacted my ways of thinking about science and mentoring. I am also grateful to all the members of the \href{https://teelabiisc.wordpress.com/}{Theoretical Ecology and Evolution lab} and the \href{https://sites.google.com/view/rohinibalakrishnanlab/home}{Animal Communication and Bioacoustics Lab} for providing a supportive and productive working environment.

I would also like to thank Prof. Sutirth Dey, my expert for this project. Sutirth provided constructive comments during the mid-year evaluation that helped me re-evaluate my project and its goals. Sutirth's brilliant course on evolution at IISER and various fascinating conversations with him and other lab members during my time as an intern at the \href{https://sites.google.com/a/acads.iiserpune.ac.in/sdlab/pbl-iiser-p}{Population Biology Lab} have been instrumental in shaping my thoughts on evolution.

I am grateful to many friends who made the IISER journey enjoyable. Countless long days and nights with Adithyan, Abhishek, Arjun, Chebi, Gaurav, Harshit, Manas, Milie, and Shruthi, have greatly helped me grow as a person, both intellectually and emotionally. I love all of y'all. Arjun has always been a great sounding board for my thoughts, no matter how inane, and has been very patient and supportive in more ways than I could list here. Milie is the occasional fitness friend that everyone needs, and occasionally texting her encouraged me to hit the gym at least semi-regularly even during hectic parts of the thesis project when schedules were uncertain. I am indebted to our D\&D group in general and Manas in particular for introducing me to D\&D and Critical Role, both of which were very helpful in getting through the pandemic as well as the thesis. Shruthi is a constant source of cat related mental health boosts and has more generally taught me a lot about a lot during my time at IISER. I am also grateful to Abhishek for letting me bother him with various physics-related questions throughout the project year. I entered this project as someone who knew some mathematics and some biology but quickly realized I would benefit from reading physics papers (typically written in physics language). Abhishek patiently answered all my basic questions about physics language/tools and skillfully helped me understand how physicists think. Insightful conversations about evolution with Gaurav were formative in gaining fresh perspectives on various topics. Closer to home, Akshay and Prerana have been my closest friends since high school and continue to provide unerring and unconditional love and support. I don't know where I would be without them today.

I am also grateful to various non-human companions, canine, feline, and more exotic. The dogs and cats of ISI are more helpful to my mental health than I could possibly articulate. Even though my thesis work may seem dreadfully abstract, running after pretty animals in the Western Ghats and other wild places of India is how I even got interested in science in the first place, and I certainly wouldn't be here today if the animals weren't so fascinating :)

%I would like to thank the KVPY fellowship program (now the INSPIRE program) by the Department of Science and Technology, Government of India, for providing me with financial support for the duration of my time at IISER. I would also like to thank the PARAM-Brahma facility at IISER Pune, part of the National Supercomputing Mission, Government of India, for providing me with computing resources, though I did not finally end up using the facility for this thesis.

Finally, I am eternally indebted to my parents, Raja and Vidya. They have quite literally supported me at every step of my life, and have nurtured and encouraged an inquisitive wonder of the world that no doubt played a vital role in helping me choose this career path. Not many parents would raise their children on Durrell and Feynman or take time out of their day to help their child find the right plant or jar for keeping an insect in the house. Upon finding out that their child has an early fascination with praying mantids and finding that there is no accessible information about the topic, not many parents would encourage their child to try and write their own source on the subject by keeping these insects at home and documenting their behavior. Upon finding out that their child was interested in snakes, not many parents would encourage this interest and actively accompany them to animal rescue centers, snake parks, and herpetology summer camps. Not many parents would be okay with their child spending weeks in the rainforest  with virtual strangers and no cellular network, catching snakes and looking for frogs in the dark with a flashlight. Certainly, not many Indian parents would let their child pursue this love of animals professionally and allow them to virtually ignore the more `mainstream' entrance exams for engineering and medical colleges in favor of pursuing the relative pipe dream of getting a KVPY scholarship for admission to an IISER or IISc. Not many parents would provide an actively supporting environment at home for an MS project, without questioning their child's future plans or PhD application process at any stage, and just trusting that the child will figure things out. Not many parents would provide fantastic support during difficult emotional situations or not blink an eye when their child decides to grow their hair out or get a tattoo (or two). Thank you for always believing in me and for letting me pursue my passions; I would be nothing without you. 
}
\allowdisplaybreaks %allow align envts to go across pages

%%%%%%%%%%%%%%%%%%%
%  The Document   %
%%%%%%%%%%%%%%%%%%%

\begin{document}
\microtypesetup{disable}
\pagestyle{plain} %Don't make the initial frontmatter fancy

%Add the frontmatter
\thesisfront

%wrap the declaration of every new \part in the \clear and the group to ensure that the fresh page created by \part does not have a page number.
% This hack-ey workaround is from https://tex.stackexchange.com/a/191189 and is needed because \part defines \pagestyle WITHIN the definition itself and so you cannot use \thispagestyle to mess with it
\cleardoublepage
\begingroup
\makeatletter
\let\ps@plain\ps@empty
\part{Motivation \& Outline}
\endgroup

%Enable microtype from here on out
\microtypesetup{enable}

%Implement the fancy headers for everything from here on out
\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{\thepage}
\fancyhead[CE]{\rule[-4ex]{0pt}{4ex}\footnotesize\itshape Eco-evolutionary dynamics of finite populations from first principles}
\fancyhead[CO]{\rule[-4ex]{0pt}{4ex}\footnotesize\itshape\nouppercase{\leftmark}}
\setlength{\headheight}{40pt} % as requested by fancyhdf
\renewcommand{\headrulewidth}{0pt}

\renewcommand{\chaptermark}[1]{\markboth{#1}{}} %gets rid of chapter number from header
\renewcommand{\sectionmark}[1]{\markright{#1}} %gets rid of section number from header

%From here on, we just need to import TeX files from the ind_files folder to include the actual content of the thesis

%Intro
\chapter{Introduction}
\input{ind_files/01_Introduction}

%The 'meat': Methods, results, theory, etc.
\cleardoublepage
\begingroup
\makeatletter
\let\ps@plain\ps@empty
\part{Theory}\label{part_theory}
\endgroup

\chapter{Mathematical background and an expository example}\label{chap_math_background}
\input{ind_files/02_math_background}

\chapter{Population dynamics from stochastic first principles}\label{chap_BD}
\input{ind_files/03_BD_processes}

\chapter{Stochastic field equations for the evolution of quantitative traits}\label{chap_infD_processes}
\input{ind_files/04_field_eqns}

%Discussion and summary
\cleardoublepage
\begingroup
\makeatletter
\let\ps@plain\ps@empty
\part{Major Takeaways \& Discussion}\label{part_summary}
\endgroup

\chapter{A unified view of population dynamics}\label{chap_unification}
\input{ind_files/05_unification_discussion}

\chapter{Discussion and Outlook}\label{sec_disc}
\input{ind_files/06_nontechnical_discussion}

%Add a new page after all the chapters are done, just to separate the main text from the appendices. If adding a new page results in you being on an even page, add another new page, just to ensure that you always begin with text on an odd page (useful if you intend to print the thesis at some point).
\clearpage
\null\thispagestyle{plain}\newpage
\ifodd\value{page}\else\null\thispagestyle{plain}\newpage\fi

%The \phantomsection is needed for hyperref to jump to the correct location
\cleardoublepage
\phantomsection
\appendix
\addcontentsline{toc}{part}{Appendices}

%make sure the title of the appendices say 'Appendix' and not 'Chapter' even though we declare them at `chapter level' internally
\renewcommand{\chaptername}{Appendix}

%For the headers of appendices (on odd pages), preface the appendix title with 'Appendix X', where X is the Appendix index (A,B,C,...)
\fancyhead[CO]{  \rule[-4ex]{0pt}{4ex}\footnotesize\itshape{\chaptername~\thechapter: \nouppercase{\leftmark}}}

%Import the Appendices
\chapter{From It\^{o} to Fokker-Planck}\label{App_SDE_FPE}
\input{ind_files/99_Appendix_A}
\chapter{Deriving stochastic trait frequency dynamics using It\^{o}'s formula}\label{App_density_to_freq}
\input{ind_files/99_Appendix_B}
\chapter{A Price-like equation for the variance of a type-level quantity}\label{App_stoch_var_eqns}
\input{ind_files/99_Appendix_C}
\chapter{Some Examples}\label{App_examples}
\input{ind_files/99_Appendix_D_examples}

%Bibliography

%You dont need the `Appendix X' prefix in the headers of the bibliography. 
\fancyhead[CO]{\rule[-4ex]{0pt}{4ex}\footnotesize\itshape\nouppercase{\leftmark}}
%Add the phantomsection so that hyperref knows where to go
\cleardoublepage
\phantomsection
\Huge\printbibheading
%Add an entry in the table of contents
\addcontentsline{toc}{part}{References}
%Putting \printbibheading up here makes sure the heading is printed before I change the font size below. This should also be outside of multicol if you use a multi-column bibliography.
%Change the font for the bibliography
\renewcommand*{\bibfont}{\footnotesize}	
%Typeset the bibliography little differently using ragged2e
\begingroup
\RaggedRight
%Add the bibliography
\printbibliography[heading=none]
\endgroup
%And we're done :)
\end{document}