### 2023-04-08 backend/contributions.tex
Change contributions page to conform to IISER guidelines

%----------------------------------------------
% Acknowledgements Header
\begin{center}
{\huge \bfseries Contributions}\\[1.5cm]
\end{center}
%----------------------------------------------

\setlength{\tabcolsep}{12pt} %space between the columns of the table
\begin{table}[ht!]
%\caption{}
%\begin{center}
	\begin{tabularx}{\textwidth}{ll}%   
\hline
{\bf Contributor Name} & {\bf Contributor Role} \\%
\hline
Ananda Shikhara Bhat, Vishwesha Guttal   & Conceptualization \\% Ideas; formulation or evolution of overarching research goals and aims
Ananda Shikhara Bhat                     & Methodology \\% Development or design of methodology; creation of models
Ananda Shikhara Bhat                     & Software \\% Programming, software development; designing computer programs; implementation of the computer code and supporting algorithms; testing of existing code components.
Ananda Shikhara Bhat, Vishwesha Guttal   & Validation \\%	Verification, whether as a part of the activity or separate, of the overall replication/reproducibility of results/experiments and other research outputs.
Ananda Shikhara Bhat                     & Formal analysis \\% Application of statistical, mathematical, computational, or other formal techniques to analyze or synthesize study data.
Ananda Shikhara Bhat                     & Investigation \\% Conducting a research and investigation process, specifically performing the experiments, or data/evidence collection.
Vishwesha Guttal                  	     & Resources \\% Provision of study materials, reagents, materials, patients, laboratory samples, animals, instrumentation, computing resources, or other analysis tools.
Ananda Shikhara Bhat                     & Data Curation \\%Management activities to annotate (produce metadata), scrub data and maintain research data (including software code, where it is necessary for interpreting the data itself) for initial use and later reuse.
Ananda Shikhara Bhat                     & Writing - Original Draft \\% Creation and/or presentation of the reported work, specifically writing the initial draft (including substantive translation).
Ananda Shikhara Bhat, Vishwesha Guttal   & Writing - Review and Editing \\% Preparation, creation and/or presentation of the reported work by those from the original research group, specifically critical review, commentary or revision - including pre- or post-acceptance stages.
Ananda Shikhara Bhat                     & Visualization \\% Preparation, creation and/or presentation of the reported work, specifically visualization/data presentation.
Rohini Balakrishnan, Vishwesha Guttal    & Supervision \\%	Oversight and leadership responsibility for the research activity planning and execution, including mentorship external to the core team.
Ananda Shikhara Bhat, Vishwesha Guttal 	 & Project Administration \\%Management and coordination responsibility for the research activity planning and execution.
---										& Funding Acquisition \\%Acquisition of the financial support for the project leading to this publication.
\hline
\end{tabularx}
%\end{center}
\label{contributions_table}
\end{table}%
This contributor syntax is based on the Journal of Cell Science CRediT Taxonomy\footnote{\href{https://journals.biologists.com/jcs/pages/author-contributions}{https://journals.biologists.com/jcs/pages/author-contributions}}